Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-14 paper_jasa.tex
tweakign language, working on section 6.5

\section{Effect of Lineup Size and Number of Observers}

@@

We need to make two crucial assumptions:
\begin{enumerate}
\item  the plot setup is structured in a way that makes it possible for observer to identify a deviation from the null hypothesis,
\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
\end{enumerate}
We will be able to show evidence for the second assumption with the help of the study discussed in Section~\ref{sec:simulation}, \hh{the degree to which the first assumption is fulfilled is reflected} by the power of a lineup. The better suited a design is for a particular task, the higher its power will be.

@@

This section takes the data and looks at the assumption made for the early expected power.

Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in table \ref{tbl:pval_assumption} for both experiment 1 and 2. The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}. We see the same pattern in figure \ref{fig:P-val_rank} where the median number of people who pick the data plot with minimum $p$-value (rank 1) are much higher than other ranks.

@@

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{p_val_rank_counts12.pdf}}
       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1). This indicates that most of the subjects pick the plot with minimum p-value (rank 1). There are subjects choosing the lineups with highest p-value(rank 20) but we can attribute this behavior to the type-I error. Can we?}
       \label{fig:P-val_rank}
\end{figure*}

\green{I think it is very interesting to look at the individual lineups by beta, and other parameters, and we can see the tendency to pick the lowest $p$-value. Figure \ref{fig:P-val_log}.}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{p_val_log_counts.pdf}}
       \caption{Frequency of plot picks by log$_10$ $p$-value in all of the sixty (60) individual lineups for experiment 1. The colored lines indicate the $p$-value of observed data plot.}
       \label{fig:P-val_log}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.5}{\includegraphics{p_val_log_counts2.pdf}}
       \caption{Frequency of plot picks by log$_10$ $p$-value in all of the seventy (70) individual lineups for experiment 2. The colored lines indicate the $p$-value of observed data plot.}
       \label{fig:P-val_log2}
\end{figure*}

>>>>

\section{Effect of Lineup Size and Number of Observers}~\label{sec:size}

@@

We need to make two crucial assumptions:
\begin{enumerate} \itemsep 0in
\item  the plot setup is structured in a way that makes it possible for observer to identify a deviation from the null hypothesis,
\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
\end{enumerate}
We will be able to show evidence for the second assumption with the help of the study discussed in Section~\ref{sec:simulation}, \hh{the degree to which the first assumption is fulfilled is reflected} by the power of a lineup. The better suited a design is for a particular task, the higher its power will be.

@@

In Section \ref{sec:size} in order to make comparisons with classical tests an assumption was made that the observer would pick the plot in the lineup that had the lowest $p$-value. In this section we take a look at the picks that subjects made, to see if this is, indeed, true. Figure \ref{fig:P-val_rank} shows side-by-side boxplots of rank of plot (in its lineup) and how many subjects selected the plot. \green{Maybe we don't need this plot.}

Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2} show the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively. In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizntally and the count of the number of subjects who selected this plot is displayed vertically. The dot represents this point, and the lines connect vertically with 0 count. Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest. Columns correspond to the other levels of the exeriment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate. (Empty cells indicate no lineup shown with this combination.) Red indicates the actual data plot, which is not always the plot in the lineup with the lowest $p$-value. As $\beta$ gets larger (stronger effect) people tended to select the plot with the lowest $p$-value.

%Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in table \ref{tbl:pval_assumption} for both experiment 1 and 2. The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}. We see the same pattern in figure \ref{fig:P-val_rank} where the median number of people who pick the data plot with minimum $p$-value (rank 1) are much higher than other ranks.

@@

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{p_val_rank_counts12.pdf}}
       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1). This indicates that most of the subjects pick the plot with minimum p-value (rank 1). There are subjects choosing the lineups with highest p-value(rank 20) but we can attribute this behavior to the type-I error. Can we?  \green{Maybe don't need this plot}}
       \label{fig:P-val_rank}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{p_val_log_counts.pdf}}
       \caption{Frequency of plot picks by log$_{10}$ $p$-value in all of the sixty (60) individual lineups for experiment 1. The colored lines indicate the $p$-value of actual data plot.}
       \label{fig:P-val_log}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.5}{\includegraphics{p_val_log_counts2.pdf}}
       \caption{Frequency of plot picks by log$_{10}$ $p$-value in all of the seventy (70) individual lineups for experiment 2. The colored lines indicate the $p$-value of actual data plot.}
       \label{fig:P-val_log2}
\end{figure*}
